---
title: "final_analysis"
author: "Liz Clark"
date: "2022-12-15"
output: html_document
---



# Part 2: Data Analysis
<style>
  .nav-pills>li>a:hover, .nav-pills>li>a:focus, .nav-pills>li.active>a,     .nav-pills>li.active>a:hover, .nav-pills>li.active>a:focus{
    background-color: #269067;
  }
</style>
## Numerical exploration
```{r}
require("here")
delomys = read.csv(here("data", "delomys.csv"))
summary(delomys$body_mass)

summary(delomys$body_length)

shapiro.test(delomys$body_mass)

shapiro.test(delomys$body_length)

```
## Graphical exploration
```{r fig.width=10}
require(palmerpenguins)
plot(bill_length_mm ~ body_mass_g, data = penguins)
```
##A scatterplot of body mass and body length
```{r fig.width=5}
plot(body_mass ~ body_length, data = delomys)
```
##A histogram of body mass
```{r fig.width=5}
boxplot(delomys$body_mass)
```
##A histogram of body length
```{r fig.width=5}
hist(delomys$body_length)
```
##A conditional boxplot of body mass, conditioned on species (column binomial)
```{r fig.width=10}
boxplot(formula= body_mass ~binomial,
        data= delomys)
```
##A conditional boxplot of body mass, conditioned on sex (column sex)
```{r fig.width=10}
boxplot(formula= body_mass ~sex,
        data= delomys)
```
##A conditional boxplot of body mass, conditioned on both species and sex

```{r fig.width=10}
boxplot(formula= body_mass ~binomial:sex,
        data= delomys)
```

## Model 1: simple linear regression body_length ~ body_mass
```{r}
fit1=lm(body_length ~ body_mass, data = delomys)
summary(fit1)
shapiro.test(fit1$residuals)
hist(residuals(fit1))

knitr::kable(coef(summary(fit1)),digits = 4)
knitr::kable(anova(fit1),digits = 4)
```
## Model 2: 1-way ANOVA body_mass ~ sex
```{r}
fit2=lm(body_mass ~ sex, data = delomys)
summary(fit2)
anova(fit2)
shapiro.test(fit2$residuals)
hist(residuals(fit2))

knitr::kable(coef(summary(fit2)),digits = 4)
knitr::kable(anova(fit2),digits = 4)
```
## Model 3: 1-way ANOVA body_mass ~ binomial
```{r}
fit3=lm(body_mass ~ binomial, data = delomys)
summary(fit3)
anova(fit3)
hist(residuals(fit3))
shapiro.test(fit3$residuals)
knitr::kable(coef(summary(fit3)),digits = 4)
knitr::kable(anova(fit3),digits = 4)
```
## Model 4: 2-way additive ANOVA body_mass ~ sex + binomial
```{r}
fit4=lm(body_mass ~ sex+binomial, data = delomys)
summary(fit4)
anova(fit4)

hist(residuals(fit4))
shapiro.test(fit4$residuals)
knitr::kable(coef(summary(fit4)),digits = 4)
knitr::kable(anova(fit4),digits = 4)
```
## Model 5: 2-way factorial ANOVA body_mass ~ sex * binomial
```{r}
fit5=lm(body_mass ~ sex*binomial, data = delomys)
summary(fit5)
anova(fit5)

hist(residuals(fit5))
shapiro.test(fit5$residuals)
knitr::kable(coef(summary(fit5)),digits = 4)
knitr::kable(anova(fit5),digits = 4)
```
## AIC model comparison
```{r}

AIC(
  fit1,
  fit2,
  fit3,
  fit4,
  fit5)

```
